<!--
本文件用途: SingleCellExperiment / scater / scran 的入门笔记(你自己写的)
数据: GSE126030, 由文中的 curl/wget chunk 下载

⚠ 注意本文件的定位: 文件名带 _Not_good 且位于 back/ 目录, 属于被搁置的尝试稿,
  不是 53.Single_and_Spatial 的正式流程 —— 正式流程见同级目录的 3_seurat_*.R。

⚠ 其它:
  - 下载 chunk 标了 eval=FALSE, 但后面的 read_tsv 直接按【当前工作目录】读文件,
    若没先手动下载并解压, knit 会在那里失败。
  - 文中有 ?here 这类交互式帮助调用, 属草稿残留。
-->

---
title: "singleCellExperiment"
author: "Zhen Gao"
date: '2024-08-01'
output: html_document
editor_options: 
  chunk_output_type: console
chunk_output_type: console
---

### introduction to singleCellExperiment 

references:

https://www.bioconductor.org/packages/release/bioc/vignettes/SingleCellExperiment/inst/doc/intro.html


```{r}
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("SingleCellExperiment")
# BiocManager::install("scater")
# BiocManager::install("scran") # <<<============ NOTIECE POPUP!!!!

# cell ranger 
# seurat eigengene Popularized by its use in Seurat, graph-based clustering 

library(SingleCellExperiment)
library(scater)
library(scran)
```

https://ftp.ncbi.nlm.nih.gov/geo/series/GSE126nnn/GSE126030/suppl/

```{bash eval=FALSE}
# I run this code with ==> Code ==> Terminal ==> send to terminal
# OR use git bash to run  curl 
# or curl # now works in windows 
curl -O https://ftp.ncbi.nlm.nih.gov/geo/series/GSE126nnn/GSE126030/suppl/GSE126030_RAW.tar
wget https://ftp.ncbi.nlm.nih.gov/geo/series/GSE126nnn/GSE126030/suppl/GSE126030_RAW.tar

tar -xvf GSE126030_RAW.tar
```


```{r}
library(tidyverse)
library(here) # here() starts at C:/Users/zhen-/Documents
?here
# here("C:/Users/zhen-/Downloads/")

counts_df<- read_tsv( ("GSM3589407_PP002swap.filtered.matrix.txt.gz"))
#counts_df<- read_tsv(here("C:/Users/zhen-/Downloads/GSM3589407_PP002swap.filtered.matrix.txt.gz"))
dim(counts_df)
counts_df[1:2,1:5]

colnames(counts_df) %>% tail()

counts_df$...2466 %>% head()

table(counts_df$...2466, useNA="ifany")
```

remove the last column
```{r}
length(colnames(counts_df))
counts_df<- counts_df[, -2466]

counts_df[1:5, 1:5]

counts_mat<- counts_df[, -c(1,2)] %>%  as.matrix()

rownames(counts_mat)<- counts_df$Gene

counts_mat[1:5, 1:5]
```

### Create a SingleCellExperiment Object 
```{r}
sce <- SingleCellExperiment(assays = list(counts = counts_mat))

sce
```
To access the count data we just supplied, we can do any one of the following:
```{r}
assay(sce, "counts")

assay(sce, "counts")[1:5, 1:5]

counts(sce)[1:5, 1:5]
```

One can access the slots with the `@` operator. This is considered bad practice as the class developers are free to alter the internal structure of the class, at which point any code using `@` may no longer work. Rather, it is best to use the provided **getter** functions like `assay()` and `counts()`

```{r}
sce@assays

sce@assays@data

sce@assays@data$counts[1:5, 1:5]
```

add more assays
```{r}
sce <- scuttle::logNormCounts(sce)
sce
```

```{r}
sce@assays@data$logcounts[1:5, 1:5]

assay(sce, "logcounts")[1:5, 1:5]

dim(logcounts(sce))
```

You may add the assay manually

```{r}
counts_1 <- counts(sce) + 1

assay(sce, "counts_1") <- counts_1 # assign a new entry to assays slot

assays(sce)

# not recommended way to add new assay data
sce@assays@data$counts_2<- counts(sce) + 2
assays(sce)
```

```{r}
assayNames(sce)
```

### add metadata

```{r}
sce <- scuttle::addPerCellQC(sce)


colData(sce)

coldata<- colData(sce)

identical(rownames(coldata), colnames(counts_mat))

sce <- scuttle::addPerFeatureQC(sce)
rowData(sce)
```

```{r}
rowRanges(sce)
```

### other metadata

the metadata slot, a named list of arbitrary objects. For example, say we have some favorite genes (e.g., highly variable genes) that we want to store inside of sce for use in our analysis at a later point. 

```{r}
my_genes <- c("gene_1", "gene_5")
metadata(sce) <- list(favorite_genes = my_genes)
metadata(sce)
```

### dimension reduction

```{r}
sce <- scater::logNormCounts(sce)

sce <- scater::runPCA(sce)

sce

reducedDim(sce, "PCA")[1:5, 1:5]

dim(reducedDim(sce, "PCA"))

sce <- scater::runTSNE(sce, perplexity = 0.1)

reducedDim(sce, "TSNE")[1:5, 1:2]

reducedDims(sce)

reducedDimNames(sce)
```

### add 

```{r}
u <- uwot::umap(t(logcounts(sce)), n_neighbors = 2)
reducedDim(sce, "UMAP_uwot") <- u

reducedDims(sce) # Now stored in the object.
```


```{r}
colLabels(sce) <- scran::clusterCells(sce, use.dimred="PCA")
table(colLabels(sce))
```

```{r}
scater::plotReducedDim(sce, dimred="TSNE")
```
