<!--
本文件用途: maftools 入门与 R Markdown 排版试验(你自己写的, 见 YAML 的 author)

⚠ 两处会影响使用:
  1) 下面有【五个 chunk 都取名 libarary】(且是 library 的拼写错误)。
     knitr 遇到重复 chunk label 会直接报 "Duplicate chunk label" 并中止编织,
     除非设置 options(knitr.duplicate.label = 'allow')。建议改成各不相同的名字。
  2) 这些 chunk 里的 install.packages / BiocManager::install 没有加 eval=FALSE,
     所以每次 knit 都会真的去联网安装, 其中 BiocManager::install("maftools", force=TRUE)
     还会强制重装。若只想留作笔记, 给这些 chunk 加 eval=FALSE。
-->

---
title: "Demon Of RMarkdown"
author: "Zhen Gao"
date: "2025-10-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Mutation Annotation Format (MAF) 
This package attempts to summarize, analyze, annotate and visualize MAF files any studies in MAF format, including TCGA 
maftools functions can be categorized into mainly Visualization and Analysis modules.

Install from Bioconductor repository 


```{r libarary}
install.packages("mclust")
```
```{r libarary}
install.packages("devtools")
```
```{r libarary}
# install.packages("Matrix") # need R 4.4 or higher to install this package 
# conda search -c conda-forge r-matrix
# conda install -c conda-forge r-matrix -y # 1.65n

```
```{r libarary}
# library(devtools)
# install_version("Matrix", version = "1.6-0", repos = "https://cran.r-project.org")
```

```{r libarary}
# Install from GitHub repository
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

BiocManager::install("maftools", force = TRUE)

```


```{r cars}
library(maftools)
```



```{r pressure, echo=FALSE}
# path to TCGA LAML MAF file
# AML as the abbreviation for acute myeloid leukemia
# LAML used within TCGA for unkown reason 
# https://gdc.cancer.gov/resources-tcga-users/tcga-code-tables/tcga-study-abbreviations

laml.maf = system.file('extdata', 'tcga_laml.maf.gz', package = 'maftools') 

#clinical information containing survival information and histology. This is optional
laml.clin = system.file('extdata', 'tcga_laml_annot.tsv', package = 'maftools') 

laml = read.maf(maf = laml.maf, clinicalData = laml.clin)

laml

```

2 space in the end of line means change line 

General Markdown Formatting (for both HTML and PDF where supported):
Bold: **bold text** or __bold text__  
Italic: *italic text* or _italic text_  
Italic bold: *__italic text__* or _**italic text**_  
Strikethrough: ~~strikethrough~~  
Code: `` code ``  
Superscript: text^super^  
Subscript: text~sub~  


<span style="color: red;">This text is red.</span>  
<span style="color: blue;">This text is blue.</span>  
<span style="color: orange;">This text is blue.</span>  
<span style="color: purple">This text is blue.</span>  
<span style="color: lightgray">This text is blue.</span>  
<span style="color: lightblue">This text is blue.</span>  
<span style="color: limon">This text is blue.</span>  
<span style="color: cyan">This text is blue.</span>  
<span style="color: magenta">This text is blue.</span>  
<span style="color: olive">This text is blue.</span>  
<span style="color: pink">This text is blue.</span>  

<span style="color: #FF5733;">This text uses a hexadecimal color code.</span> 


foo <font color='brown'>bar</font> foo


<a >blue color</a>

	`red color`

#### list 

ordered: 

1. Main item one
    1. Sub-item one
    1. Sub-item two
2. Main item two
    1. Another sub-item


unordered 1:

* Item 1
* Item 2
  * Nested Item 2.1
  * Nested Item 2.2


unordered 2:

+ Item 3
+ Item 4
+ Item 4

unordered 1:

- Item 3
- Item 4
- Item 4
